MSE and dynamic state variable model with sigma function

### 2018-05-03 doc/MSE_MAR_18.tex
modification in tex

Management strategy evaluation (MSE; \cite{Sainsbury2000, Smith1994, Bunnefeld2011}) seeks to study the implications of management strategies using simulation\cite{Punt2016}. Such simulation should include all important processes of a fishery, which is inherently a socio-ecological system, with coupled dynamics of the fishing fleets, the exploited stocks, and their governance\cite{Punt2016,Rasemeyer2007}. Given the complexity of such a socio-ecological system the uncertainy about the dynamics and feedbacks are consideable. As a result, any forecasted success in achieving goals in a management strategy may thus be inperceivably small compared to the stochasticity in results generated in the system, or worse, be counteracted by unintended consequences of the management strategy.The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.

The adaptive response of fishers to their environment is one of the key uncertainties requiring attention in MSE \cite{Fulton2007}. Simulation tools for such adaptive responses are available in fisheries, borrowing methodology from the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}. New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010, Gillis1995} have been developed. These models generally predict the effect in the short term (within a fishing trip, or a quota year) by optimizing a utility function and determine which choices best yields the best chance of increasing utility, while keeping track of the state of each individual. The effect of a choice on the utility depends on the economic environment,   such as the home port of the vessel and the distance to fishing grounds, and the bioological enviroment, such as the spatial distribution of the resources. Using such dynamic variable state models, effects of e.g. relocation costs of marine protected areas have been modelled \cite{Dowling2011}.

Dynamic state variables have also been used to describe discarding behaviour \cite{Gillis1995, Batsleer2013, Batsleer2015}. These studies may shed light on the potential outcomes of the European fisheries management reform of 2013. That reform included using the use of the Maximum Sustainable Yield (MSY) reference points as targets for exploiting commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}. Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted  high costs in mixed fisheries in the short term as a result of the LO. These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{ Alzorriz2018}. This will result in fishig effort reallocation and early closures of fisheries once quota have been reached. However, the  long term benefits to fisheries that could result from improved selectivity are ignored in those studies. Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO. Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches. Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Ulrich2017, Batsleer2013}. Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.

@@

Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels.to understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

@@

The OM captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.  The operating model thus includes  individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance. The  management body used the MP to make its decision on how to respond to the state of the resource. This management procedure thus includes the data collection from the fishery, how these are interpreted, and the Harvest Control Rule that dictates the limits that are set to the fishery.

\begin{figure}[!h]
\centering
\includegraphics[width=0.7\textwidth]{Figures/MSE.eps} 
\caption{The conceptual framework and how it was mapped.}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.6\textwidth]{Figures/Areadynamics.eps} 
\caption{The conceptual framework and how it was mapped.}
\label{fig:MSE}
\end{figure}

@@

\begin{equation} \label{eq:solve}
P_{p, d}(C_i, Q_i, F_i, s) = \frac
                % Nominator
                {e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}}
                % Denominator
                {\sum_p \sum_d e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}},
\end{equation}
where $\sigma$ was a tuning parameter that measures how important was to be near optimal. A high value of sigma resulted in uniform probabilities of choices, and that vessels were distributed uniformly among the different fishing areas. In contrast, if sigma was small then all vessels concentrated in the optimal location (but note that $\sigma$ should be $> 0$). For computations, we used $\sigma = 40$ (noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).

@@

 (six seasons) and spatially (two areas) explicit
The maximum age in the age-structured fish model was 4 years.
The two stocks are named "stock 1" and "stock 2". and the areas are named "area a" and "area b".
In the example used for our simulations, it was assumed that there were 100 new-borns per year for both stocks. The difference between the stocks is that all new-borns of stock 1 were born in "area a", while all new-borns of stock 2 were born in "area b".

>>>>

Management strategy evaluation (MSE; \cite{Bunnefeld2011, Sainsbury2000, Smith1994}) seeks to study the implications of management strategies using simulation\cite{Punt2016}. Such simulation should include all important processes of a fishery, which is inherently a socio-ecological system, with coupled dynamics of the fishing fleets, the exploited stocks, and their governance\cite{Punt2016,Rasemeyer2007}. Given the complexity of such a socio-ecological system the uncertainy about the dynamics and feedbacks are consideable. As a result, any forecasted success in achieving goals in a management strategy may thus be inperceivably small compared to the stochasticity in results generated in the system, or worse, be counteracted by unintended consequences of the management strategy. The robustness of management strategies to the uncertainy in the processes that govern fisheries systems thus need to be accounted for \cite{Andersen2010, Kell2007, Prellezo2016, Punt2016}.

The adaptive response of fishers to their environment is one of the key uncertainties requiring attention in MSE \cite{Fulton2007}. Simulation tools for such adaptive responses are available in fisheries, borrowing methodology from the more general question of state-dependent foraging decisions in ecological systems \cite{ClarkandMangel2000,Houston1999}. New tools for state-dependent behaviour of individual fishing vessels, translated into behaviour of the fleet and implemented using stochastic dynamic programming \cite{Alzorriz2018,Batsleer2015, Dowling2011, Poos2010, Gillis1995} have been developed. These models generally predict the effect in the short term (within a fishing trip, or a quota year) by optimizing a utility function and determine which choices best yields the best chance of increasing utility, while keeping track of the state of each individual. The effect of a choice on the utility depends on the economic environment, such as the home port of the vessel and the distance to fishing grounds, and the biological enviroment, such as the spatial distribution of the resources. Using such dynamic variable state models, effects of e.g. relocation costs of marine protected areas have been modelled \cite{Dowling2011}.

Dynamic state variables have also been used to describe discarding behaviour \cite{Batsleer2015, Batsleer2013, Gillis1995}. These studies may shed light on the potential outcomes of the European fisheries management reform of 2013. That reform included using the use of the Maximum Sustainable Yield (MSY) reference points as targets for exploiting commericially important fish stocks and the gradual introduction of a landings obligation (LO) \cite{CFP2013}. Indeed, using spatial and temporal distributions of catch rates within a single year, dynamic state variable models forecasted  high costs in mixed fisheries in the short term as a result of the LO. These costs result from removing the option to adapt landings to quota by discarding parts of the catch \cite{Alzorriz2018}. This will result in fishig effort reallocation and early closures of fisheries once quota have been reached. However, the  long term benefits to fisheries that could result from improved selectivity are ignored in those studies. Ignoring the potential for improved selectivity leaves out a key element in the perceived benefit of the LO. Because the LO should provide incentives for the use of more selective gears and for fisheries to  move away from areas with high levels of unwanted catch \cite{Alzorriz2016, Alzorriz2018}, the improved selectivity should lead to higher long term catches. Achieving single species MSY while landing all commercial catches in fisheries targeting multiple species (mixed fisheries) is challenging because achieving the objective for one species may mean missing the objective for another \cite{Batsleer2013,Ulrich2017}. Forecasting whether such management will be effective depends on understanding the response of individual effort allocation to the management.

@@

Using a MSE, the main objective of this work is not to reproduce the entire complexity of a real mixed fishery, but to defined plausible hypothesis about population dynamics and then to implemented the processes of interest, i.e., changes in population abundance, to finally measure the effect of the potential consequences associated with the fishers optimal choice location underlying assumption of the current state-dependent behaviour of individual fishing vessels. To understand the fleet dynamics in a spatially and temporally heterogeneous mixed quota-regulated fishery when reducing from unmanaged (unconstrained catch quota) to MSY managed and to compare it with a situation where discards are allowed in order to assess on the biological and economic consequences. If LO and catch quotas are not fully enforced, there will be an incentive to continue discarding. Therefore, this manuscript tends to understand the likely adaptive change in fishing patterns that will help us to understand if this leads to a better balance between quotas and catches.

@@

The OM captured the key processes in the dynamics of the fish populations, the fisheries and the management body, and can be thought of as a minimum realistic model \cite{Punt1995}.  The operating model thus includes  individual harvester decision-making (including error) and the consequent biomass of fish stocks, including the essential elements for calculating the individual harvester economic performance. The  management body used the MP to make its decision on how to respond to the state of the resource  (Figure \ref{fig:stockdyn}). This management procedure thus includes the data collection from the fishery, how these are interpreted, and the Harvest Control Rule that dictates the limits that are set to the fishery.

\begin{figure}[!h]
\centering
\includegraphics[width=0.55\textwidth]{Figures/MSE.eps} 
\caption{The Managament Strategy Evaluation framework including the Operating Model (OM) and the Management Procedure (MP).}
\label{fig:MSE}
\end{figure}

\begin{figure}[!h]
\centering
\includegraphics[width=0.55\textwidth]{Figures/Areadynamics.eps} 
\caption{The Operating Model (OM) including the key processes in the dynamics of the fish populations and how it was mapped.}
\label{fig:stockdyn}
\end{figure}

@@

\begin{equation} \label{eq:solve}
P_{p, d}(C_i, Q_i, F_i, s) = \frac
                % Nominator
                {e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}}
                % Denominator
                {\sum_p \sum_d e^{ -\Delta_{p, d}(C_i, Q_i, F_i, s)/\sigma}},
\end{equation}
where $\sigma$ was a tuning parameter that measured how important it was to be near the optimal choice. A large $\sigma$ resulted in uniform probabilities of choices, with vessels being distributed uniformly across the different fishing areas. In contrast, a small $\sigma$ forces vessels to concentrate in the optimal location (but note that $\sigma$ should be $> 0$). For computations, we used $\sigma = 40$ (noting that $\Delta$ ranged from 0 to $\sim 2 x 10^{3}$, but was generally of the order of $1 x 10^2$ in magnitude).

@@

 (six seasons) and spatially (two areas) explicit
The maximum age in the age-structured fish model was 4 years.
The two stocks are named ``stock 1" and ``stock 2". and the areas are named ``area a" and ``area b". 
In the example used for our simulations, it was assumed that there were 100 new-borns per year for both stocks. The difference between the stocks is that all new-borns of stock 1 were born in ``area a", while all new-borns of stock 2 were born in ``area b".
